Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-05-19 doc/report/msc-report.tex
doc/report/msc-report.tex: * implementation chapter.

In the first section the implementation of project creation, configuration, and
setup, will be presented. The second section features the data import
implementation. The third section presents the implementation of data
interchange between Zohmg and other software but also throughout the Zohmg
system. The fourth, and final, section presents the implementation details of
the Zohmg system's exporting compontent.

@@

A standard MapReduce map function emits key-value pairs. Not so in the case of
the Zohmg mapper: the usermap function emits triples (that is, a tuple of three
values) consisting of a timestamp, a list of dimensions and their respective
values, and a list of measurements. In effect, the dimensions and the timestamp
specify a point in n-space and the measurements represent the observations made
at that point.

% TODO: move apache logs to introduction perhaps.
An example data source is the Apache HTTP Server access logs. The Apache HTTP
Server Combined Log Format is commonly used to log requests. The log format is
configurable, an example log line might have the following format

@@

The above example log entry is line-broken due to space constraints, normally
the entry would be on one line. The example log line contains access info for a
specific request; it contains the key entries: the client's IP address,
timestamp, requested resource, HTTP status code for the request, the client's
user agent etc.

Revisiting the running example of Apache access logs, a usermap function might
look something like this:

@@

For every line of the log file, a point in n-space (the dimensions and the
timestamp) is emitted together with a count of pageviews and bytes; one
pageview per request and the number of bytes served in the request. It is
possible to add business logic here, perhaps only counting requests from
useragents that are not web spiders which resulted in a HTTP 200 status code.

@@

@@

\begin{verbatim}
   zohmg import mappers/apache.py \
   /data/weblogs/2009/05/10
\end{verbatim}

This will instruct zohmg to apply the map function found in \texttt{apache.py}
to all logs that it can find in the directory
\texttt{/data/weblogs/2009/05/10} on HDFS. The emitted output is duly
collected and projections are carefully stored in the data store.

% TODO: implementation.
-- The implementation of all that --

Since the wish is to store projections rather than all combinations of all
dimensions, the data needs to be mangled slightly before sending it down
the pipe to the underlying data store.

Dumbo expects a key-value pair from the mapper. Hadoop will collect all keys
and give the reducer a key and an iterator of values.

If projections are to be stored, both the desired dimensions and the unit
in question needs to be encoded in the key, and the value of the unit
emitted in the \textit{value} field.

By doing this it is possible to use a simple SumReducer to sum the values
for each unit at any point in n-space.

Remember, the usermapper gives us a point in n-space and a list of
measurements. We need to break out the measurements and sum over each such
measurement. Zohmg solves this by wrapping the usermapper in a cleverly named
devicie, namely a mapperwrapper.

% TODO: this para sounds as if it was written by a drunken monkey.
The mapperwrapper takes each tuple emitted by the usermapper and performs a
dimensionality reduction, ending up with the projections requested by the
user. For every projection and unit, the dimensions remaining after the
reduction are appended to the string representation of the unit and emitted
as the key. The value is simply the value of the unit. The Hadoop framework
will do the heavy work of collecting all keys and summing the values.

An alternative approach would have been to perform the dimensionality
reduction at a later stage, namely in the reducer of a multi-pass MapReduce
job.

The reducer, then, is very simple: it is a simple sum of the values. The output
of the reducer will not be stored on disk as is usually the case, but will
instead be interpreted by a custom OutputReader that persists the data to
HBase.

@@

% TODO: data interchange.
The reducer needs to communicate to the OutputReader under what rowkey,
column-family and qualifier to store the computed sum. It does this by
emitting the row key as the key and a JSON-serialized map of the form
\texttt{\{"column-family:qualifier" : value\}}.

@@

TableInputFormat

TextInputFormat

LzoInputFormat

@@

% TODO: not truly correct
The Darling output format for HBase. Between Dumbo, which Zohmg uses for data
import, and Hadoop JSON is used for serializing data.

The Darling output format increased import speed by a factor ten, compared to
using Thrift.

@@

The transformers application fetches data from HBase with the same mechanism as
the raw data application. The data is transformed with a transformer---a Python
program in the project's \texttt{transformers} directory---requested by the
user. Similar to the raw data serving, the output is served as JSONP.

@@

Summary here.

@@

Fetches data from HBase and transforms it with the requested transform
program from the project's \texttt{transformers} directory, output is
dumped as JSONP.

@@

\noindent The above query extracts the same data as the query from the Data
section. This data is run through the \texttt{transformer.py} application and
is, again, served to the user as JSONP.

>>>>

In the first section the implementation of project creation, configuration, and
setup, will be presented. The second section features the data import
implementation; it will present example data as well as an example user map
function to import this. The third section presents the implementation of data
interchange between Zohmg and other software, such as Hadoop, Dumbo, and HBase,
but also throughout the Zohmg system. The third section also presents the
implementation details of the Zohmg system's exporting compontent.

@@

A standard MapReduce map function emits key-value pairs. The Zohmg usermap
function, instead, emits triples (tuples with three values), which consist of a
timestamp, a hash table of dimensions and their respective values, and a hash
table with measurements (values). In effect, the dimensions and the timestamp
specify a point in n-space and the measurements represent the observations made
at that point.

% TODO: move apache logs to introduction perhaps.
An example data source is the Apache HTTP Server access logs. The Combined Log
Format is commonly used to log Apache HTTP Server requests. The log format is
configurable, an example log line might have the following format

@@

The above example log entry is line broken due to space constraints, normally
the entire entry would be on a single line. The example log line contains access
info for a specific request; it contains the key entries: the client's IP
address, timestamp, requested resource, HTTP status code for the request, the
client's user agent etc.

Processing---which means: read, analyze, and store---logs of the format
mentioned in Listing 5.1, requires the user to write a map function. This user
map function might look something like

@@

For every line of the log file (the value argument to the map function), a point
in n-space (the dimensions and the timestamp) is emitted together with a count
of pageviews and bytes; one pageview per request and the number of bytes served
in the request. It is possible to add logic here, perhaps only counting requests
from useragents that are not web spiders which resulted in a HTTP 200 status
code.

@@

Since the wish is to store projections rather than all combinations of all
dimensions, the data needs to be mangled slightly before sending it down
the pipe to the underlying data store. Dumbo expects a key-value pair from the
mapper. Hadoop will collect all keys and give the reducer a key and an iterator
of values.

% TODO: ok to remove this?
%If projections are to be stored, both the desired dimensions and the unit
%in question needs to be encoded in the key, and the value of the unit
%emitted in the \textit{value} field.
%
%By doing this it is possible to use a simple SumReducer to sum the values
%for each unit at any point in n-space.

Remember, the usermapper gives us a point in n-space and a list of
measurements. We need to break out the measurements and sum over each such
measurement. Zohmg solves this by wrapping the user map function.

The wrapper takes each tuple emitted by the user map function and performs a
dimensionality reduction, ending up with the projections requested by the user.
For every projection and unit, the dimensions remaining after the reduction are
appended to the string representation of the unit and emitted as the key. The
value is simply the value of the unit. The Hadoop framework will do the heavy
work of collecting all keys and summing the values.

% TODO: Move this to DISCUSSION.
%An alternative approach would have been to perform the dimensionality
%reduction at a later stage, namely in the reducer of a multi-pass MapReduce
%job.

The reducer, then, is very simple: it is a simple sum of the values. The output
of the reducer will not be stored on disk as is usually the case, but will
instead be interpreted by a custom OutputReader (see Data Interchange section)
that persists the data to HBase.

@@

\begin{verbatim}
   zohmg import mappers/apache.py \
   /data/weblogs/2009/04/20
\end{verbatim}

This will instruct zohmg to apply the map function found in \texttt{apache.py}
to all logs that it can find in the directory \texttt{/data/weblogs/2009/04/20}
on HDFS. The emitted output is duly collected and projections are carefully
stored in the data store.

@@

% TODO: data interchange.
The reducer needs to communicate to the OutputReader under what rowkey,
column-family and qualifier to store the computed sum. It does this by
emitting the row key as the key and a JSON-serialized map of the form \\
\texttt{\{"column-family:qualifier" : value\}}.

@@

The simplest input format would be the TextInputFormat, this is used to parse
flat text files which reside on HDFS.

Hadoop is able to read LZO compressed files with the LzoInputFormat. In order to
be able to do so effectively the compressed files first have to be indexed with
a tool that is provided with Hadoop.

% TODO: When is the TableInputFormat used.
The TableInputFormat is able read input to a Hadoop program from an HBase table.

@@

% TODO: not truly correct
% TODO: A detailed explanation of generally everything in the darling dir.
The Darling output format for HBase. Between Dumbo, which Zohmg uses for data
import, and Hadoop JSON is used for serializing data.

The Darling output format increased import speed by a factor ten, compared to
using Thrift. The HBase Thrift interface is slow because it first spawns a Java
HBase client which interprets everything from the HBase Thrift server.
This actually adds an extra step to the interaction with HBase making it slower
than talking to it natively with Java.

@@

The transformer middleware application fetches data from HBase with the same
mechanism as the raw data server (via the Thrift interface); it then uses a user
transformer to alter the data and return it to the user. The user transformer is
a Python program in the project's \texttt{transformers} directory. See the
appendix A for more information.

The data is received in the transformer application as a hash table, which is
then available for manipulation. The result is returned from the transformer
application as a hash table and the transformer middleware returns the
transformed result to the user as JSONP.

@@

This chapter's first section has presented the implementation of the Zohmg
system. The concept of Zohmg project's and management of thesame were
introduced; as well as how these are configured. The second section presented
how to import data with Zohmg, giving examples of data and the user map
function. In the third, and final, section Zohmg's data interchange was
featured. Both the internal data interchange, which is used together with
Hadoop, Dumbo, and HBase, and the external data interchange, which is used by
the user to query Zohmg.

@@

It is possible to transform the raw data in Zohmg by using a so called
\textit{transformer}. The transformer application receives a query---of the same
format as the raw data server's---and then transforms the data with the
requested transformer program from the project's \texttt{transformers}
directory, output is dumped as JSONP.

@@

The above query extracts the same data as the query from the Data
section. This data is run through the \texttt{transformer.py} application and
is, again, served to the user as JSONP.
